---
title: "OCLC API Project"
author: "Joseph Branson"
date: '2019-02-27'
output: html_document
slug: oclc-api
tags:
  - book project
  - portfolio
  - programming
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Background
I'm a bit of a compulsive book collector. It started slowly, but through the years I've accrued a few volumes. Generally, I like to associate each book with the circumstances under which I received/acquired it - these were for a paper about Frantz Fanon, these were after visiting Europe and becoming really interested in modern architecture, etc.

Previously, this was also my method for organization. Books were grouped together by origin story. Sometimes this led to thematic groupings, but other times there appeared no apparent connection.

Alas, after a few changes in apartment this system broke down, until I was left with an *almost* random ordering. Devising a system for reorganization on my own seemed an impossible task (or you know, really hard). However, there are other systems for such organization - namely, the Dewey Decimal Classification (DDC) and the Library of Congress Classification (LCC). 

##Purpose
This post is the result of my efforts to obtain DDC and LCC information for the associated ISBN values using the [OCLC Classify API](http://classify.oclc.org/classify2/api_docs/index.html). It is also my first significant endeavor in using R - I am very much looking for any code review or feedback.

##Process
```{r message=FALSE}
library(xml2)
library(plyr)
library(tidyverse)
library(DT)
```

```{r message=FALSE}
book_csv <- "library_20190223200451.csv"
book_list <- read_csv(book_csv)
isbn_list <- book_list$isbn13
options(scipen = 99)
```

I used [Libib](https://www.libib.com) to scan the ISBN barcodes from the books when available or input manually when not. It populated quite a bit of info itself.

```{r echo=FALSE}
glimpse(book_list)
datatable(head(book_list[ ,c("authors","title", "publish_date", "pages","isbn13")]), rownames = FALSE)
```


My general approach was to break the process of retrieving one book's classification into steps and then `map()` that over the whole list of books. 

  1. Get the XML object from OCLC
  2. Parse that object and extract DDC, LCC, and OWI
  3. Attach classification info back to respective book


### Get the XML object from OCLC
This function takes a value and a classification system (currently only "isbn13" and "owi"), queries the OCLC API, and returns a list (from xml) object for that value. 

Future: generalize `type`, add support for other API fields, toggle for summary
```{r}
oclc_get_raw <- function(number, type="isbn13") {
  
    base_OCLC <- "http://classify.oclc.org/classify2/Classify?"
    end_base_OCLC <- "&summary=true"

  if(tolower(type) == "isbn13") {
    lookup_type <- "isbn"
  } else if(tolower(type) == "owi") {
    lookup_type <- "owi"
  } else {
    message("Classification type not recognized.")
    break
  }
  call_address <- paste0(base_OCLC, lookup_type, "=", number, end_base_OCLC)
  oclc_xml_list <- read_xml(call_address) %>% as_list()
  
}
```

This next function is a little out of order, but I think I have to define it before it gets called in the following function. It simply retrieves a value ("ddc", "lcc", or "owi") from one of the xml_list objects and returns NA if not found. Some entries were missing DDC or LCC classifications. If they were missing OWI values, all three values would be missing.

My lesson: R won't return NA if you point to nonexistent values; it will return NULL, which isn't a vector or list, thus causing further issues.
```{r}
retrieve_val <- function(value, xml_list) {
  x1 <- attributes(xml_list[["classify"]][["recommendations"]][[value]][["mostPopular"]])[["nsfa"]]
    ifelse(!is.null(x1), x1, NA)
}

#Hopefully this fixes the 1d thing, might've been missing some OWI
retrieve_owi <- function(value, xml_list) {
  x1 <- attributes(xml_list[["classify"]][["work"]])[[value]]
    ifelse(!is.null(x1), x1, NA)
}
```

### Parse that object and extract DDC, LCC, and OWI
Each xml object had 1 of 3 codes:
- "0" indicates a successful query
- "4" indicates multiple entries for that ISBN value (not possible for OWI)
- "102" indicates no entry found

Below I treated 102 as an error and re-queried.

For each "4", I retrieved the OWI of the first listing and used it to submit another query. Each OWI could only have a single result.

After receiving an xml-list object with code "0", I extracted the classifications using the above functions and put them together in a data frame, replacing NULL values with NA.

One of the messiest things about this code is all the repeated brackets and `attributes()`. I suspect there's a way to clean that up with `%>%` and `~.` but I'll have to investigate further.

```{r}
recs_xml_list <- function(oclc_xml_list) {
  #Checking response code. "4" indicates multiple entries found for isbn, "102" indicates technical malfunction, "0" indicates successful retrieval
  code <- attributes(oclc_xml_list[["classify"]][["response"]])[["code"]]  
  
  
 if(code == 4 | code == 102) {
   #Find OWI of first work and get xml object with it instead (or just trying again in case of 102)
    owi2 <- attributes(oclc_xml_list[["classify"]][["works"]][[1]])[["owi"]]
    oclc_xml_list <- oclc_get_raw(owi2, type = "owi")
  } else  if(code==0) {
    oclc_xml_list <- oclc_xml_list
  } else {
    #included for safety?
    message("Unrecognized error code.")
    break
  }
    #If successful retrieval, some books won't have lcc or dcc. These receive NA for the respective value.  
      ddc <- retrieve_val("ddc", oclc_xml_list)
      lcc <- retrieve_val("lcc", oclc_xml_list)
      owi <- retrieve_owi("owi", oclc_xml_list)
  #I was getting an error here about owi being a 1d vector or list so I am messing around with it
  #Edit: turns out it was because some ISBN values aren't recorded, so OWI was returning NULL, which isn't 1d vector or list
  recs_df <- data_frame(ddc = ddc, lcc = lcc, owi = owi)
  recs_df
}
```


### Attach classification info back to respective book
This code runs over the whole `isbn13` column in `book_list`. The first command is what takes the longest. I'm not sure what the limiting factor is there, but speeding it up would definitely be nice.

I also don't love that I've made it stick together a whole list of xml objects then parse them. I suspect it would be faster to do one query, parse, return values, then move on to next query.

Future: I want to make a way to attach values back to original list by matching the ISBN. Right now it just collects them all into a long data frame and glues them back to the side.
```{r cache=TRUE, results="hold"}
raw_list <- map(book_list$isbn13, oclc_get_raw) 
rec_list <- map(raw_list, recs_xml_list)

classification <- ldply(rec_list, rbind)

sum(is.na(classification$ddc))

sum(is.na(classification$lcc))

sum(is.na(classification$owi))

nrow(classification)
```

```{r}
book_list <- cbind(book_list, classification)

isbn_class <- cbind(isbn = book_list$isbn13, classification)

```

```{r echo=FALSE}
datatable(book_list[ ,c("authors","title", "publish_date", "pages","isbn13", "ddc","lcc", "owi")], rownames = FALSE, options = list(
  columnDefs = list(list(className = 'dt-center', targets = 6)),
  pageLength = 40,
  lengthMenu = c(15, 40, 60, 100),
  scrollX=TRUE
))
```


```{r eval=FALSE}
ggplot(isbn_class, aes(x=isbn, y = owi)) +
  geom_jitter()
```
##Lessons
This final approach is the product of many dead ends. My initial googling turned up [Open Library Books API](https://openlibrary.org/dev/docs/api/books), which looked very promising. It could send JSON objects, and it had tons of information about each title. Unfortunately, it was missing quite a few entries (~60).

I then found the OCLC API, but after struggling with the less-than-cooperative XML formats I resumed searching to find the [ISBNdb API](https://isbndb.com/apidocs). Although it requires a paid plan, I conceded that a paid service ($10) would probably be more complete and responsive. The cost would certainly be worth a few hours of hacking.

In the end, it also wasn't as complete as I would hope. However, I did learn how to submit `GET` requests using `add_headers()`. My lesson: if you want `X-API-KEY: MY-KEY`, the function will add ": " itself - `add_headers("X-API-KEY" = my_key)`.

```{r}
#Need function to update df individually, not in bulk. Also search based on matching isbn, not position in list
amend_list <- function(isbn, type, large_df) {
  out_list <- oclc_get_raw(isbn) %>% recs_xml_list()
  large_df[which(large_df$isbn13==isbn), type] <- out_list[[type]]
            }




```